Landau quantization of nearly degenerate bands, \\ \red{and non-crossing rules for Landau levels}

### 2018-12-23 paper.tex
Updates from ShareLaTeX

\documentclass[aps, prb, showpacs, twocolumn, notitlepage, superscriptaddress]{revtex4-1}
\usepackage{amsmath}
\usepackage{amssymb,amscd,bm,dsfont,wasysym,mathrsfs,latexsym,psfrag,accents,mathtools}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{color}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{boldline}
\usepackage[normalem]{ulem}
\input{preamble}

\renewcommand{\topfraction}{1}
\renewcommand{\bottomfraction}{1}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{1}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{3}

@@

\title{Landau quantization of nearly degenerate bands}

\author{Chong Wang}\affiliation{Institute for Advanced Study, Tsinghua University, Beijing 100084, China}
\author{Wenhui Duan}\affiliation{Institute for Advanced Study, Tsinghua University, Beijing 100084, China}\affiliation{Department of Physics and State Key Laboratory of Low-Dimensional Quantum Physics, Tsinghua University, Beijing 100084, China}\affiliation{Collaborative Innovation Center of Quantum Matter, Tsinghua University, Beijing 100084, China}
\author{Leonid Glazman}\affiliation{
Department of Physics, Yale University, New Haven, Connecticut 06520, USA}
\author{A. Alexandradinata}\affiliation{
Department of Physics, Yale University, New Haven, Connecticut 06520, USA}

\begin{abstract}
In the semiclassical theory of Landau levels, a Bloch electron is driven by a magnetic field  along constant-energy band contours  known as cyclotron orbits. We introduce a quantization rule for Landau levels of nearly-degenerate orbits whose differential area ($\delta S$) in $\bk$-space is much smaller than their average area $S$. The splitting of two nearly-degenerate orbits commonly originates from spin-orbit coupling, or from the breaking of a crystalline point-group symmetry. By treating both $\delta S/S$ and $1/l^2S$ ($l$ being the magnetic length) as small parameters with finite ratio $l^2\delta S$, our perspective diverges from and generalizes previous semiclassical theories.  $1/l^2\delta S$ measures the hybridization of split orbits due to the Zeeman interaction; in the opposing regimes $l^2\delta S {{\ll}} 1$  and $l^2\delta S{\gg}1$, we recover the known Onsager-Lifshitz-Roth rules for nondegenerate and exactly-degenerate orbits, respectively.  In the absence of crystalline point-group symmetries, three tunable real parameters are needed to attain a spin-degeneracy between two Landau levels; we have exhaustively identified all symmetry classes of orbits for which this number is reduced from three. In particular, only one parameter is needed for rotational-symmetric orbits; this parameter may be the magnitude or orientation of the field, or the bias voltage in tunneling spectroscopy. A signature of single-parameter tunability is a smooth crossover between period-doubled and -undoubled quantum oscillations in the low-temperature, Schubnikov-de Haas effect. We demonstrate the utility of our quantization rule, as well as the tunability of Landau-level degeneracies, for the Rashba two-dimensional electron gas -- subject to an arbitrarily oriented magnetic field, and possibly augmented by the Dresselhaus spin-orbit interaction.
\end{abstract}
\date{\today}

@@

@@

Semiclassical quantization rules are well-known for the Landau levels of  both nondegenerate\cite{Onsager,lifshitz_kosevich,lifshitz_kosevich_jetp} and exactly-degenerate\cite{rotheffham,rothmag,topoferm,100p,Mikitik_quantizationrule} energy bands; the latter is exemplified by the spin-degenerate bands of centrosymmetric solids.
These rules become increasingly accurate as the area of the cyclotron orbit becomes much larger than the inverse square of the magnetic length $l{=}\sqrt{\hbar/eB}$.

Our work generalizes these rules to the Landau quantization of nearly-degenerate bands, whose cyclotron orbits are split in $\bk$-space with a differential area  $\delta S$ that is much less than their average area $S$; $\delta S/S{\ll}1$ shall be referred to as the \textit{near-degeneracy} condition. Nearly-degenerate orbits commonly originate from the spin-orbit-splitting of energy bands, as exemplified by a two dimensional electron gas (2DEG) with Rashba spin-orbit coupling:
\begin{equation}
H_R=\frac{{\hbar^2} k^2}{2m}+\hbar\alpha  (k_{x}\sigma_{y}-k_{y}\sigma_{x}).\label{eq:Rashba-Hamiltonian}
\end{equation}
As illustrated in Fig.\ \ref{fig:orbits}(a), the near-degeneracy condition is satisfied at energies much greater than $m\alpha^2$:\footnote{The expressions for $\delta S$ and $S$ in \q{deltaSvsS} are asymptotically valid for $E{\gg}m\alpha^2$.}
\e{\delta S = 4\pi m \alpha k_E/\hbar\ll S =\pi k_E^2; \;\; k_{E}:=\f{\sqrt{2m E}}{\hbar}.\label{deltaSvsS}}
$S$ is equivalently the area of the  constant-energy contour in the absence of spin-orbit coupling ($\alpha{=}0$).

@@

By treating both $\delta S/S$ and $1/l^2S$  as small parameters with finite ratio $l^2\delta S$, our perspective diverges from and generalizes previous semiclassical theories\cite{kohn_effham,blount_effham,rotheffham,wannier_fredkin,gao_zero-field_2017,topoferm,100p,Mikitik_quantizationrule} which have only a single, small parameter ($1/l^2S$). $1/l^2\delta S$ measures the hybridization of nearly-degenerate orbits by the Zeeman interaction.
In the regime $l^2\delta S{\gg}1$, the Onsager-Lifshitz quantization rules apply independently to  two concentric (but decoupled) orbits. In the opposite regime $l^2\delta S{\ll}1$, the spin-orbit splitting of energies is negligible, and we may then apply the known quantization rule for exactly-degenerate bands\cite{rotheffham,rothmag,topoferm,100p,Mikitik_quantizationrule}. The goal of this work is to formulate a unifying rule that interpolates smoothly between these two known regimes; such a rule is presented in \s{sec:qtznrules}.

In \s{sec:Rashba}, we demonstrate the utility of this rule for the Rashba 2DEG with an arbitrarily-oriented magnetic field, and possibly augmented by the Dresselhaus spin-orbit coupling. A sufficiently strong in-plane field not only moves the Dirac point but tilts the Dirac cone over -- into a type-II Dirac point\cite{soluyanov_type-ii_2015}. In \s{sec:inplanezeeman}, we investigate the Landau-Zener tunneling dynamics that occur in the vicinity of a II-Dirac point; in particular, we will critically evaluate a recent claim\cite{obrien_magnetic_2016} of perfect Klein tunneling at the energy of the Dirac point.

A symmetry analysis (cf.\ \s{sec:llquasideg}) of our quantization rule allows us to determine how many tunable parameters are needed to attain a (pseudo-)spin degeneracy between two Landau levels. This degeneracy is exact in symmetry classes where Landau levels are distinguishable by different quantum numbers; absent such quantum numbers, it is still possible for the repulsion between two Landau levels to be anomalously small -- this notion of a Landau-level quasidegeneracy is formulated in \s{sec:relatedegeneracies}. Due to a rotational symmetry of the Rashba-Dresselhaus 2DEG, we find that only a single tunable parameter is needed to attain a Landau-level quasidegeneracy (cf.\ \s{sec:ctsrot}); this parameter can be the magnitude or orientation of the field, or the bias voltage in tunneling spectroscopy.

@@

%is not traceless, the traceful part can always be absorbed into the redefinition of $S$ within the accuracy of first order perturbation theory.

%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

%the wavepacket;  $g_s{\approx}2$ is the free-electron $g$-factor and $\mu_B$ the Bohr magneton.[100page] 
%Spin in $z$ direction is represented by $s_z=\hbar\sigma_z/2$ \textbf{[detail]}.

@@

\qq{eq:rule}{berryconn} allow us to determine $\lambda_{1,2}$ to the lowest order  in $1/l^2S$ and $\delta S/S$, and for arbitrary values of $l^2\delta S$ and $|B\calm|/\delta \var_{\pm}$. We remark that the quantization rule is invariant under basis transformations within the nearly-degenerate band subspace (see App. \ref{app:quantizationruleproof}); in subsequent discussions, it may be more convenient to adopt a $\bk$-independent basis where the non-abelian Berry connection vanishes.

@@

\noi{i} If $|\var_+{-}\var_-|$ is everywhere (along $\frako_0$) much greater than $B|\calm_{+-}|$, then an electron would undergo adiabatic (band-conserving) dynamics\cite{nenciu_review} along two independent, nondegenerate orbits. Such an adiabatic regime always exists as $B{\rightarrow} 0$ for nearly-degenerate orbits that do not intersect, e.g, at a type-II Dirac point, which is the subject of the case studies in \s{sec:inplanezeeman} and \s{sec:rotsymmbreakdown}.
By  neglecting  $B\calm_{+-}$ in favor of intraband elements,  $\cala$ reduces to a diagonal matrix with  diagonal entries ($e^{i\lambda_{\pm}}$) given by:
\e{\lambda_{\sma{\pm}} = -\int_0^{T_c}\calh_{\sma{\pm\pm}}\f{dt}{\hbar}=\pm \f{1}{2}l^2\delta S - B\int_0^{T_c}\calm_{\sma{\pm\pm}}\f{dt}{\hbar}.  \label{phaseindependentorbit}} 
\q{eq:rule}, combined with \q{phaseindependentorbit}, may be identified as the Onsager-Lifshitz rule\cite{Onsager,lifshitz_kosevich,lifshitz_kosevich_jetp} for two independent, nondegenerate orbits of areas $S{\pm}\delta S$, and with subleading phase corrections that were first identified by Roth\cite{rothmag}.\footnote{We remark that \q{phaseindependentorbit} neglects the contribution by band $-$ to the orbital moment of band $+$: $\delta M_z=il^{-2}\epsilon_{\ab}(\delta \var_+-\delta \var_-)\mathfrak{X}^{\alpha}_{+-}\mathfrak{X}^{\beta}_{-+}$. $\mathfrak{X}^{\alpha}_{+-}$, being an off-diagonal matrix element of the position operator, is generically of order a lattice constant $a$. We then estimate $\int_0^{T_c} B\delta M_z dt/\hbar=O(a^2 \delta S)$, which is negligible compared to $l^2\delta S/2$. Note also that the single-band orbital moment simply vanishes in some symmetry classes of orbits\cite{100p}.} The perturbative inclusion of $B\calm_{\sma{+-}}$ results in a correction to $\lambda_{\pm}$ [cf.\ \q{phaseindependentorbit}] that is expressible as a power series in $B$, with a leading term proportional to $1/B$; this will be confirmed in a subsequent case study [cf.\ \q{lambdasmallB}] and applied to understand beatings in quantum-oscillatory phenomena in \s{sec:qo}.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{Characterization of the Rashba 2DEG with an in-plane Zeeman field. (a) illustrates the trefoil-shaped cyclotron orbit at the energy of the II-Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the degeneracy-lifting Hamiltonian $\delta\var$ [cf.\ \q{eq:RZ-Hamiltonian}]. The wavefunction along the inner (resp.\ outer) component of the trefoil orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-splitting field. The upper panel of (b) schematically illustrates  Landau levels near the II-Dirac point (in the weak-hybridization regime). The dashed line represents the Onsager-Lifshitz quantization of the trefoil orbit [cf.\ Eq.\ (\ref{trefoilrule})], and the solid line incorporates the first-order correction [cf.\ Eq.\ (\ref{RZfirstorderE})] due to tunneling. Lower panel of (b): red dashed line shows the energetic difference between Landau levels determined by numerical diagonalization and Landau levels obtained by the trefoil quantization rule  [cf.\ Eq.\ (\ref{trefoilrule})]; this red dashed line may be compared to the perturbative formula $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})], which is plotted as a black solid line.  This comparison is made (at each $1/B$) for the Landau level closest to 0.45eV and with even Landau-level index $n$; the II-Dirac point is located at 0.5eV.  Parameters for lower panel of (b): $m{=}0.076m_0$, $\alpha{=} 1.5\times10^{5}$cm/s, $Z_\parallel {=}$1meV. Out-of-plane Zeeman coupling is negligibly weak for the choice of $m$. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. Inset: an enlarged view of near II-Dirac point. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4{<}\epsilon_0{<}1.0$(eV)\cite{RZfignote}. Parameters for (d): $m=0.076m_0$, $l=100$\AA.\label{fig:RZ}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{LL.png}
\caption{A schematic diagram of Landau level dispersion (solid lines) with respect to $B$ for nearly-degenerate orbits. $\delta\lambda$ is plotted in the colored background. Landau level quasidegeneracies occur at $\delta\lambda\equiv 0\equiv 2\pi$.\label{fig:LL}}
\end{figure}

@@

In \s{sec:relatedegeneracies}, we will describe the precise relation between eigenvalue-degeneracies of  $\A$ and (near) degeneracies of Landau levels. We then ask how many tunable real parameters are needed to attain an eigenvalue-degeneracy of $\A$. Our answer is  $3$ in the absence of symmetry, as discussed in \s{sec:introducecodimension}; the answer may reduce to $0,1$ or $2$ depending on the symmetry class of the orbit, as will be explained in \s{sec:tenfold}. Before this general symmetry classification,  we will first develop physical intuition by studying two rotationally-symmetric models [cf.\ \s{sec:singleparameterrashba} and \ref{sec:rotsymmbreakdown}] where only one tunable parameter is needed -- this can be the $B$ field itself.

@@

Since $\A$ depends on $\alpha,~\beta,~l,~E,~m$ only through $\hbar\alpha k_E /\var_c$, $\hbar\beta k_E /\var_c$ and $g_sm/m_0$, let us consider the three-dimensional parameter space $(\hbar\alpha k_{\sma{E}}/\var_c,\hbar\beta k_{\sma{E}}/\var_c,g_sm/m_0){\in} \mathbb{R}^3$.  $\A(\hbar\alpha k_{\sma{E}}/\var_c,\hbar\beta k_{\sma{E}}/\var_c,g_sm/m_0){=}{\pm} I$ determines a set of concentric circles in the $\beta{=}0$ plane, as per the $\mathfrak{c}_{\infty}$-symmetric case study in \s{sec:ctsrot}. Moving off this plane, we find that $\A{=}I$ is satisfied in a  neighborhood of $\beta{=}0$, i.e., $\A{=}I$ defines a hypersurface in $\mathbb{R}^3$ that is illustrated in \fig{fig:dgn}(a). On the other hand, $\A{=}{-}I$ is not satisfied in the neighborhood of $\beta{=}0$, hence the circles associated to $\A{=}{-}I$ remain isolated as line nodes, as illustrated in \fig{fig:dgn}(b). To recapitulate, we have found that one of every two Landau-level crossings (those labelled by odd $j$) destabilize due to the symmetry reduction.

>>>>

\documentclass[aps, showpacs, twocolumn, notitlepage, superscriptaddress]{revtex4-1}
\usepackage{amsmath}
\usepackage{amssymb,amscd,bm,dsfont,wasysym,mathrsfs,latexsym,psfrag,accents,mathtools}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{color}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{boldline}
\usepackage[normalem]{ulem}
\input{preamble}

\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{3}

@@

\title{Landau quantization of nearly degenerate bands, \\
\red{and non-crossing rules for Landau levels}}

\author{Chong Wang}\affiliation{Institute for Advanced Study, Tsinghua University, Beijing 100084, China}
\author{Wenhui Duan}\affiliation{Institute for Advanced Study, Tsinghua University, Beijing 100084, China}\affiliation{Department of Physics and State Key Laboratory of Low-Dimensional Quantum Physics, Tsinghua University, Beijing 100084, China}\affiliation{Collaborative Innovation Center of Quantum Matter, Tsinghua University, Beijing 100084, China}
\author{Leonid Glazman}\affiliation{
Department of Physics, Yale University, New Haven, Connecticut 06520, USA}
\author{A. Alexandradinata}\affiliation{
Department of Physics, Yale University, New Haven, Connecticut 06520, USA}\affiliation{Department of Physics and Institute for Condensed Matter Theory, University of Illinois at Urbana-Champaign, Urbana, Illinois 61801, USA}

\begin{abstract}
In the semiclassical theory of Landau levels, a Bloch electron is driven by a magnetic field  along constant-energy band contours  known as cyclotron orbits. We introduce a quantization rule for Landau levels of nearly-degenerate orbits whose differential area ($\delta S$) in $\bk$-space is much smaller than their average area $S$. The splitting of two nearly-degenerate orbits commonly originates from spin-orbit coupling, or from the breaking of a crystalline point-group symmetry. By treating both $|\delta S/S|$ and $1/l^2|S|$ ($l$ being the magnetic length) as small parameters with finite ratio $l^2|\delta S|$, our perspective diverges from and generalizes previous semiclassical theories.  $1/l^2|\delta S|$ measures the hybridization of split orbits due to the Zeeman interaction; in the opposing regimes $l^2|\delta S| {\ll} 1$  and $l^2|\delta S|{\gg}1$, we recover the known Onsager-Lifshitz-Roth rules for nondegenerate and exactly-degenerate orbits, respectively. We apply our quantization rule to determine the number of tunable, real parameters needed to attain a \red{(pseudo)}spin degeneracy between two Landau levels, \red{i.e., a `non-crossing rule' for Landau levels. Three parameters are needed}  in the absence of crystalline point-group symmetries; we have exhaustively identified all symmetry classes of orbits for which this number is reduced from three. In particular, only one parameter is needed for orbits which are invariant under spatial rotation \red{or inversion}; this single parameter may be the magnitude or orientation of the field, or the bias voltage in tunneling spectroscopy. Signatures of single-parameter tunability include (i) a smooth crossover between period-doubled and -undoubled quantum oscillations in the low-temperature Schubnikov-de Haas effect, as well as (ii) `magic angle' magnetoresistance oscillations. We demonstrate the utility of our quantization rule, as well as the tunability of Landau-level degeneracies, for the Rashba-Dresselhaus two-dimensional electron gas -- subject to an arbitrarily oriented magnetic field.
\end{abstract}
\date{\today}

@@

%n close analogy with the Wigner-von Neumann `non-crossing rule' in quantum mechanics, we find that three tunable parameters are needed to attain a Landau-level quasidegeneracy -- in the absence of crystalline point-group symmetries [cf.\ \s{sec:introducecodimension}]. 
%We have further determined symmetric variants of the ‘non-crossing rule’ for Landau levels, for any spacetime symmetry that occurs in crystalline solids

@@

\begin{itemize}
    \item \red{fix all $|S|$ and $\delta S$}
\end{itemize}

\red{Semiclassical quantization rules compactly describe the energy dispersion of Landau levels, and are predictive of quantum oscillations in transport\cite{SdH} and thermodynamic\cite{dHvA} phenomena. The original formulation of the rule  -- by Onsager\cite{Onsager} and Lifshitz\cite{lifshitz_kosevich,lifshitz_kosevich_jetp} -- neglects the effect of spin-orbit coupling. While subsequent generalizations by Roth
\cite{rotheffham,rothmag}
and the present authors\cite{topoferm,100p} account for the spin-orbit effect, such quantization rules are limited in application to either a single nondegenerate energy band, or to energy-degenerate bands (as exemplified by spin-degenerate bands in centrosymmetric solids). It is now recognized that the spin-orbit effect is encoded in subleading corrections to the quantization rule –-- which are associated to the Berry phase and the orbital magnetic moment.}

\red{The Onsager-Lifshitz-Roth rule remains  unapplicable to a wide variety of spin-orbit-coupled metals lacking a spatial inversion center, or having magnetic order. The energy bands of such metals carry a spin degree of freedom, and the spin degeneracy of energy bands is weakly lifted by spin-orbit coupling. A prototypical example is given by  a two dimensional electron gas (2DEG) with Rashba spin-orbit coupling:}
\begin{equation}
H_R=\frac{{\hbar^2} k^2}{2m}+\hbar\alpha  (k_{x}\sigma_{y}-k_{y}\sigma_{x}).\label{eq:Rashba-Hamiltonian}
\end{equation}
\red{Given a magnetic field in the $-z$ direction, and focusing on energies much greater than $m\alpha^2$, the cyclotron orbits [illustrated in \fig{fig:orbits}(a)]   are split in $\bk$-space with a differential area  $\delta S$ that is much less than their average area $S$:}\footnote{The expressions for $\delta S$ and $S$ in \q{deltaSvsS} are asymptotically valid for $E{\gg}m\alpha^2$.}
\e{\delta S = 4\pi m \alpha k_E/\hbar\ll S =\pi k_E^2; \;\; k_{E}:=\f{\sqrt{2m E}}{\hbar}.\label{deltaSvsS}}
\red{$S$ is equivalently the area of the  constant-energy contour in the absence of spin-orbit coupling ($\alpha{=}0$).}

@@

\red{More generally, we refer to 
$\delta S/S{\ll}1$ (defined geometrically as the ratio of differential to average area) as the \textit{near-degeneracy} condition for energy bands. Our work presents a generalized rule to describe the Landau quantization of nearly-degenerate bands, for which  spin-orbit-split bands is but one example. This rule becomes increasingly accurate as the average area $S$ becomes much larger than the inverse square of the magnetic length $l{=}\sqrt{\hbar/eB}$.} By treating both $\delta S/S$ and $1/l^2S$  as small parameters with finite ratio $l^2\delta S$, our perspective diverges from and generalizes previous semiclassical theories\cite{kohn_effham,blount_effham,rotheffham,wannier_fredkin,gao_zero-field_2017,topoferm,100p,Mikitik_quantizationrule} which have only a single, small parameter ($1/l^2S$). $1/l^2\delta S$ measures the hybridization of nearly-degenerate orbits by the Zeeman interaction.
In the regime $l^2\delta S{\gg}1$, the Onsager-Lifshitz\red{-Roth} quantization rules apply independently to  two concentric (but decoupled) orbits. In the opposite regime $l^2\delta S{\ll}1$, the spin-orbit splitting of energies is negligible, and we may then apply the known quantization rule for exactly-degenerate bands\cite{rotheffham,rothmag,topoferm,100p,Mikitik_quantizationrule}. Our unifying rule provides the smooth interpolation between these two known regimes. \red{Alternatively stated, we have extended the applicability of quantization rules to the regime where the Zeeman interaction competes with the spin-orbit interaction, where previously such rules were known only when one interaction dominates over the other. Indeed, in the absence of magnetic breakdown[references],

This unifying rule is presented in \s{sec:qtznrules} for nearly-degenerate bands with a (pseudo)spin degree of freedom. As elaborated in \s{sec:discussion}, the most general formulation of our rule is applicable  to any number of nearly-degenerate bands of any symmetry.}

In \s{sec:Rashba}, we demonstrate the utility of this rule for the Rashba\red{-Dresselhaus} 2DEG with an arbitrarily-oriented magnetic field. A sufficiently strong in-plane field not only moves the Dirac point but tilts the Dirac cone over -- into a type-II\cite{soluyanov_type-ii_2015, muechler_tilted_2016, bergholtz_topology_2015} Dirac point. In \s{sec:inplanezeeman}, we investigate the Landau-Zener tunneling dynamics that occur in the vicinity of a II-Dirac point; in particular, we will critically evaluate a recent claim\cite{obrien_magnetic_2016} of perfect Klein tunneling at the energy of the Dirac point.

\red{\s{sec:llquasideg} applies} our quantization rule to determine how many real, tunable parameters are needed to attain a \red{(pseudo)}spin degeneracy between Landau levels. \red{This degeneracy may be exact if Landau-quantized wave functions transform in different unitary representations of a spatial symmetry; absent such a symmetry,} it is still possible for the repulsion between two Landau levels to be anomalously weak -- this notion of a Landau-level quasidegeneracy is formulated in \s{sec:relatedegeneracies}. \red{In close analogy with the Wigner-von Neumann `non-crossing rule' in quantum mechanics\cite{neumann2000behaviour}, we find that three tunable parameters are needed to attain a Landau-level quasidegeneracy -- in the absence of crystalline point-group symmetries [cf.\ \s{sec:introducecodimension}]. 
We have further determined symmetric variants of the Landau-level ‘non-crossing rule’, for any spacetime symmetry that occurs in crystalline solids; our results are summarized in \tab{table:codimension}. In particular, we highlight that in the presence of a spatial rotation or inversion symmetry,}  only a single tunable parameter is needed to attain a Landau-level quasidegeneracy (cf.\ \s{sec:ctsrot}); this single parameter may be the magnitude or orientation of the field, or the bias voltage in tunneling spectroscopy.

@@

@@

\qq{eq:rule}{berryconn} allow us to determine $\lambda_{1,2}$ to the lowest order  in $1/l^2S$ and $\delta S/S$, and for arbitrary values of $l^2\delta S$ and $|B\calm|/\delta \var_{\pm}$. We remark that the quantization rule is invariant under basis transformations within the nearly-degenerate band subspace (see App. \ref{app:quantizationruleproof}); in subsequent case studies, it may be more convenient to adopt a $\bk$-independent basis (if such a basis exists) where the non-abelian Berry connection vanishes.

@@

\noi{i} If $|\var_+{-}\var_-|$ is everywhere (along $\frako_0$) much greater than $B|\calm_{+-}|$, then an electron would undergo adiabatic (band-conserving) dynamics\cite{nenciu_review} along two independent, nondegenerate orbits. Such an adiabatic regime always exists as $B{\rightarrow} 0$ for nearly-degenerate orbits that do not intersect, e.g, at a type-II Dirac point, which is the subject of the case studies in \s{sec:inplanezeeman} and \s{sec:rotsymmbreakdown}.
By  neglecting  $B\calm_{+-}$ in favor of intraband elements,  $\cala$ reduces to a diagonal matrix with  diagonal entries ($e^{i\lambda_{\pm}}$) given by:
\e{\lambda_{\sma{\pm}} = -\int_0^{T_c}\calh_{\sma{\pm\pm}}\f{dt}{\hbar}=\pm \f{1}{2}l^2\delta S - B\int_0^{T_c}\calm_{\sma{\pm\pm}}\f{dt}{\hbar}.  \label{phaseindependentorbit}} 
\q{eq:rule}, combined with \q{phaseindependentorbit}, may be identified as the Onsager-Lifshitz rule\cite{Onsager,lifshitz_kosevich,lifshitz_kosevich_jetp} for two independent, nondegenerate orbits of areas $S{\pm}\delta S$, and with subleading phase corrections that were first identified by Roth\cite{rothmag}\footnote{We remark that \q{phaseindependentorbit} neglects the contribution by band $-$ to the orbital moment of band $+$: $\delta M_z=il^{-2}\epsilon_{\ab}(\delta \var_+-\delta \var_-)\mathfrak{X}^{\alpha}_{+-}\mathfrak{X}^{\beta}_{-+}$. $\mathfrak{X}^{\alpha}_{+-}$, being an off-diagonal matrix element of the position operator, is generically of order a lattice constant $a$. We then estimate $\int_0^{T_c} B\delta M_z dt/\hbar=O(a^2 \delta S)$, which is negligible compared to $l^2\delta S/2$. Note also that the single-band orbital moment simply vanishes in some symmetry classes of orbits\cite{100p}.}. The perturbative inclusion of $B\calm_{\sma{+-}}$ results in a correction to $\lambda_{\pm}$ [cf.\ \q{phaseindependentorbit}] that is expressible as a power series in $B$, with a leading term proportional to $1/B$; this will be confirmed in a subsequent case study [cf.\ \q{lambdasmallB}] and applied to understand beatings in quantum-oscillatory phenomena in \s{sec:qo}.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{Characterization of the Rashba 2DEG with an in-plane Zeeman field. (a) illustrates the trefoil-shaped cyclotron orbit at the energy of the II-Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the degeneracy-lifting Hamiltonian $\delta\var$ [cf.\ \q{eq:RZ-Hamiltonian}]. The wavefunction along the inner (resp.\ outer) component of the trefoil orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-splitting field. The upper panel of (b) schematically illustrates  Landau levels near the II-Dirac point (in the weak-hybridization regime). The dashed line represents the Onsager-Lifshitz-Roth quantization of the trefoil orbit [cf.\ Eq.\ (\ref{trefoilrule})], and the solid line incorporates the first-order correction [cf.\ Eq.\ (\ref{RZfirstorderE})] due to tunneling. Lower panel of (b): red dashed line shows the energetic difference between Landau levels determined by numerical diagonalization and Landau levels obtained by the trefoil quantization rule  [cf.\ Eq.\ (\ref{trefoilrule})]; this red dashed line may be compared to the perturbative formula $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})], which is plotted as a black solid line.  This comparison is made (at each $1/B$) for the Landau level closest to 0.45eV and with even Landau-level index $n$; the II-Dirac point is located at 0.5eV.  Parameters for lower panel of (b): $m{=}0.076m_0$, $\alpha{=} 1.5\times10^{5}$cm/s, $Z_\parallel {=}$1meV. Out-of-plane Zeeman coupling is negligibly weak for the choice of $m$. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. Inset: an enlarged view of near II-Dirac point. (d) For various parametrizations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in the left panel the quantity $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ [as defined in Eq. (\ref{whereisdiracpoint})] $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. The corresponding $l^2 \delta S^2/S$ is plotted in the right panel. For numerical convergence,\cite{RZfignote} we have presented data only in a strip [within ($\alpha$, $Z_\parallel$)-space] where $0.4{<}\epsilon_0{<}1.0$(eV). Parameters for (d): $m=0.076m_0$, $l=100$\AA.\label{fig:RZ}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{LL.png}
\caption{A schematic diagram of Landau level dispersion (solid lines) with respect to $B$ for nearly-degenerate orbits. $\delta\lambda$ is plotted in the colored background. Landau level quasidegeneracies occur at $\delta\lambda\equiv 0\equiv 2\pi$ (\red{indicated by red dashed lines}).\label{fig:LL}}
\end{figure}

@@

In \s{sec:relatedegeneracies}, we will describe the precise relation between eigenvalue-degeneracies of  $\A$ and (near) degeneracies of Landau levels. We then ask how many tunable real parameters are needed to attain an eigenvalue-degeneracy of $\A$. Our answer is  $3$ in the absence of symmetry, as discussed in \s{sec:introducecodimension}; the answer may reduce to $0,1$ or $2$ depending on the symmetry class of the orbit, as will be explained in \s{sec:tenfold}. Before this general symmetry classification,  we will first develop physical intuition by studying two rotationally-symmetric models [cf.\ \s{sec:singleparameterrashba} and \ref{sec:rotsymmbreakdown}] where only one tunable parameter is needed -- this can be \red{the magnitude of} the $B$ field itself.

@@

\red{For the Rashba-Dresselhaus 2DEG, $\A$ depends on $\alpha,~\beta,~l,~E,~m$ only through three independent parameters:}%$\hbar\alpha k_E /\var_c$, $\hbar\beta k_E /\var_c$ and $g_sm/m_0$, let us consider the three-dimensional parameter space:
\e{\left( \f{\hbar\alpha k_{\sma{E}}}{\var_c},\f{\hbar\beta k_{\sma{E}}}{\var_c},g_s\f{m}{m_0} \right){\in} \mathbb{R}^3. \la{parameter2}}  
$\A(\hbar\alpha k_{\sma{E}}/\var_c,\hbar\beta k_{\sma{E}}/\var_c,g_sm/m_0){=}{\pm} I$ then determines a set of concentric circles in the $\beta{=}0$ plane, as per the $\mathfrak{c}_{\infty}$-symmetric case study in \s{sec:ctsrot}. Moving off this plane, we find that $\A{=}I$ is satisfied in a  neighborhood of $\beta{=}0$, i.e., $\A{=}I$ defines a hypersurface in $\mathbb{R}^3$ that is illustrated in \fig{fig:dgn}(a). On the other hand, $\A{=}{-}I$ is not satisfied in the neighborhood of $\beta{=}0$, hence the circles associated to $\A{=}{-}I$ remain isolated as line nodes, as illustrated in \fig{fig:dgn}(b). To recapitulate, we have found that one of every two Landau-level crossings (those labelled by odd $j$) destabilize due to the symmetry reduction.
